---
title: "Figures"
author: ""
date: "4/27/2021"
output:
  pdf_document:
    latex_engine: xelatex
sansfont: Arial
mainfont: Arial
header-includes:
- \AtBeginDocument{\let\maketitle\relax}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, dev = "png", dpi = 200)
```

```{r, include=F}
library(tidyverse)
library(patchwork)
library(ggtext)
library(magick)

extrafont::loadfonts(quiet=TRUE)
```

\newpage

## Figure 1

```{r, fig.dim=c(10,6), out.height="4.2in", out.width="7in"}
read_rds("results/panels/figure1.rds")
```

\newpage

## Figure 2

```{r}
read_rds("results/panels/figure2.rds")
```

\newpage

## Figure 3

```{r, fig.dim=c(10,6), out.height="4.5in", out.width="6.5in",fig.align="center"}
read_rds("results/panels/figure3.rds")
```



\newpage

## Figure 4

```{r, fig.dim=c(10,10), out.height="7in", out.width="7in"}

read_rds("results/panels/figure4.rds")
```

\newpage

## Figure 5

```{r, fig.dim=c(14,8), out.height="4in", out.width="7in"}
read_rds("results/panels/figure5.rds")
```

\newpage

## Supplementary Figure 1

```{r, fig.dim=c(20,20), out.height="6.5in", out.width="6.5in"}
read_rds("results/panels/supp1.rds")
```

\newpage

## Supplementary Figure 2

```{r, fig.dim=c(20,20), out.height="7in", out.width="7in",fig.align='center'}
read_rds("results/panels/supp2.rds")
```

\newpage

## Supplementary Figure 3

```{r, fig.dim=c(8,10), out.height="6.4in", out.width="8in",fig.align='center'}
read_rds("results/panels/supp3.rds")
```

\newpage


## Supplementary Figure 4

```{r, fig.dim=c(8,8), out.height="7in", out.width="7in"}
read_rds("results/panels/supp4.rds")
```

## Supplementary Figure 5


```{r, fig.dim=c(8,4), out.height="5in", out.width="5in",fig.align='center'}
read_rds( "results/panels/supp4_5.rds")
```

\newpage

## Supplementary Figure 6

```{r, fig.dim=c(3.5,4), out.height="4in", out.width="7in", fig.align="center", dpi=72, dev="png"}
image_read('resources/210215_accord2_calfluor610_eachm_quasar670_3p4.slices_1_1.split_channels.png',strip = T, density = 200) %>% image_ggplot()
```


\newpage

## Supplementary Figure 7

```{r, fig.dim=c(3.5,4), out.height="5in", out.width="7in", fig.align='center', dev="png"}
image_read('resources/210215_accord2_calfluor610_eachm_quasar670_3p4.slices_1_30.montage.slices_1_30.montage.png',strip = T,density = 200) %>% image_ggplot()
```

\newpage

## Supplementary Figure 8

```{r, fig.dim=c(3.5,4), out.height="5in", out.width="7in", fig.align='center', dev="png"}
s7a <- image_read('resources/210204_dapi_quasimodo2_calfluor61_eachm_quasar670p.slices_1_1.representative.png',strip = T,density = 200) %>% image_ggplot()

#s7b <- image_read('resources/210204_dapi_quasimodo2_calfluor61_eachm_quasar670p.slices_1_22_x2.montage.png',strip = T,density = 200) %>% image_ggplot()

s7a

```

\newpage

## Supplementary Figure 9

```{r, fig.dim=c(10,12), out.height="7in", out.width="5.81in"}
read_rds("results/panels/supp5.rds")
```

\newpage

## Supplementary Figure 10

```{r, fig.dim=c(10,7), out.height="3in", out.width="7in", align="center", fig.align='center'}
read_rds("results/figs/arriba_total_rna_fusions/arriba_total_rna_fusions.ggp.rds")
```

